---
title: "Mod 4 Hw"
author: " "
output:
  pdf_document: default
  html_document: default
---

Load in the `election2012.RData` file. 

 Information on data. County level elections results
 - Totalvotes: total number of votes in the county
 - Demvotes: total number of votes received by Dem in that county
 - DemVotesTotalPercent:the percentage of the votes for the Dems in that county

```{r}
# load what libraries you think you'll need


# load data
load("election_2012.RData") 
```

1. Create a code chunk that tells us the number of rows, cols, the names of all of the variables in the data set, and the State(s) of the first few observations. What do you see? 

```{r}


```

2. What percentage of the vote did the Democratic Candidate for President in 2012 (Barack Obama)?

3. How many counties (observations/rows) does Tennessee have? 

```{r}



```

4. Visualize total votes data using a histogram. Show your code and interpret the result. Do you think this is the best way to visualize this data? Why or why not?

```{r}



```

5. Visualize the percentage of democrat votes by total vote count using a scatterplot. What can you learn from this graphic?

```{r}


```

# Wine Data

```{r}
library(readr)
winequality_red <- read_delim("winequality-red.csv", ";")
head(winequality_red)
```

6. Take a moment to check the dimensions, structure, and layout of the data. In 1-3 sentences how would you describe the data to an audience? 

```{r}

```

Below, I've overlayed a density plot onto a histogram. Investigate the code.

```{r}
ggplot(winequality_red, aes(x = alcohol)) +
      geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
      geom_density(alpha = .2, fill="#FF6655") #overlay with a transparent (alpha value) density
```

7. Next, interpret and describe the distribution of the data in complete sentences. 

Additional lines can be added to represent the “mean” using geom_vline. Can you add a vertical line representing the mean to the plot above? Show your code here: 

```{r}


```

8. In class we saw how to add the correlation coefficient to a graphic. This last part of this assignment asks you to explore, on your own, the correlation matrix, or a 'Correlogram.'

I recommend installing and playing with the package "ggcorrplot." Read the help file. Look up the package's vignette online (https://cran.r-project.org/web/packages/ggcorrplot/readme/README.html). Quickly replicate the vignette's LAST example using the built in mtcars data but change the # of digits printed to 2 not 3.  Which two variables have the highest positive correlation?

```{r}




```

